---
title: 'Practical Machine Learning Week 3 Quiz'
author: "Claus Bo Hansen"
date: "December 21, 2018"
output: ioslides_presentation
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = FALSE)

library(caret)
library(Hmisc)
library(ggplot2)
library(rpart)
library(rpart.plot)
library(rattle)
library(pgmm)

```

## Question 1

Load the cell segmentation data from the AppliedPredictiveModeling package using the commands:

1. Subset the data to a training set and testing set based on the Case variable in the data set.

2. Set the seed to 125 and fit a CART model with the rpart method using all predictor variables and default caret settings.

3. In the final model what would be the final model prediction for cases with the following variable values:

```{r, echo = TRUE}
library(AppliedPredictiveModeling)
data(segmentationOriginal)
library(caret)

# Create training set
trainingelems <- createDataPartition(segmentationOriginal$Case, list = FALSE)
trainingdata <- segmentationOriginal[trainingelems,]

set.seed((125))

# Train model
model <- rpart(Class ~., data = trainingdata)

# Plot decision tree
fancyRpartPlot(model)


```


Based on the decision tree, the model prediction is

a PS

b WS

c PS

d Not possible to predict

## Question 2

If K is small in a K-fold cross validation is the bias in the estimate of out-of-sample (test set) accuracy smaller or bigger? If K is small is the variance in the estimate of out-of-sample (test set) accuracy smaller or bigger. Is K large or small in leave one out cross validation?

The bias is larger and the variance is smaller. Under leave one out cross validation K is equal to the sample size.

## Question 3

These data contain information on 572 different Italian olive oils from multiple regions in Italy. Fit a classification tree where Area is the outcome variable. Then predict the value of area for the following data frame using the tree command with all defaults

```{r, echo = TRUE}

library(pgmm)
data(olive)
olive = olive[,-1]

model <- train(Area ~., data = olive, method = "rpart")

newdata = as.data.frame(t(colMeans(olive)))

predict(model, newdata)

```

2.783. It is strange because Area should be a qualitative variable - but tree is reporting the average value of Area as a numeric variable in the leaf predicted for newdata




## Question 4


```{r, echo = TRUE}

library(ElemStatLearn)
data(SAheart)
set.seed(8484)
train = sample(1:dim(SAheart)[1],size=dim(SAheart)[1]/2,replace=F)
trainSA = SAheart[train,]
testSA = SAheart[-train,]

set.seed(13234)

model <- train(chd ~ age+alcohol+obesity+tobacco+typea+ldl, data=testSA, method="glm", family="binomial")

predictionTest <- predict(model, testSA)
valuesTest <- testSA$chd

predictionTrain <- predict(model, trainSA)
valuesTrain <- trainSA$chd

missClass = function(values,prediction){sum(((prediction > 0.5)*1) != values)/length(values)}

# Test set
missClass(valuesTest, predictionTest)

# Train set
missClass(valuesTrain, predictionTrain)

```


Test Set Misclassification: 0.27

Training Set: 0.31

=>

Test Set Misclassification: 0.31

Training Set: 0.27

## Question 5

Load the vowel.train and vowel.test data sets:

Set the variable y to be a factor variable in both the training and test set. Then set the seed to 33833. Fit a random forest predictor relating the factor variable y to the remaining variables. Read about variable importance in random forests here: http://www.stat.berkeley.edu/~breiman/RandomForests/cc_home.htm#ooberr The caret package uses by default the Gini importance.

Calculate the variable importance using the varImp function in the caret package. What is the order of variable importance?

```{r, echo = TRUE}

library(ElemStatLearn)
data(vowel.train)
data(vowel.test)

vowel.test$y <- as.factor(vowel.test$y)
vowel.train$y <- as.factor(vowel.train$y)

set.seed(33833)

model <- train(y ~ ., data = vowel.train, method = "rf")

varImp(model)

```

rf variable importance

     Overall
x.2  100.000
x.1   97.511
x.5   39.196
x.6   25.532
x.8   22.663
x.4    9.684
x.3    7.975
x.9    5.655
x.7    3.589
x.10   0.000

