---
title: 'Week 2: The Caret Package Quiz'
author: "Claus Bo Hansen"
date: "November 20, 2018"
output: ioslides_presentation
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = FALSE)

library(caret)
library(Hmisc)
library(ggplot2)

```

## Question 1

Load the Alzheimer's disease data using the commands:

```{r, echo = TRUE}
library(AppliedPredictiveModeling)
data(AlzheimerDisease)

```


Which of the following commands will create non-overlapping training and test sets with about 50% of the observations assigned to each? 

```{r, echo = TRUE}

# Option 1
adData = data.frame(diagnosis,predictors)
trainIndex = createDataPartition(diagnosis, p = 0.50)
#training = adData[trainIndex,]
#testing = adData[-trainIndex,]
# Returns error: Error in `[.default`(xj, i) : invalid subscript type 'list'

# Option 2
adData = data.frame(diagnosis,predictors)
train = createDataPartition(diagnosis, p = 0.50,list=FALSE)
test = createDataPartition(diagnosis, p = 0.50,list=FALSE)
# Creates overlapping sets

# Option 3
adData = data.frame(diagnosis,predictors)
trainIndex = createDataPartition(diagnosis, p = 0.50,list=FALSE)
training = adData[trainIndex,]
testing = adData[-trainIndex,]
# Correctly creates non-overlapping sets

# Option 4
adData = data.frame(predictors)
trainIndex = createDataPartition(diagnosis,p=0.5,list=FALSE)
training = adData[trainIndex,]
testing = adData[-trainIndex,]
# Is missing diagnosis

```


## Question 2

Load the cement data using the commands:

```{r, echo = TRUE}
library(AppliedPredictiveModeling)
data(concrete)
library(caret)
set.seed(1000)
inTrain = createDataPartition(mixtures$CompressiveStrength, p = 3/4)[[1]]
training = mixtures[ inTrain,]
testing = mixtures[-inTrain,]

```


## Question 2, featurePlot


```{r, echo = TRUE}

featurePlot(x=training[,c("Cement", "BlastFurnaceSlag","FlyAsh", "Water", "Superplasticizer", "CoarseAggregate", "FineAggregate", "Age")], y = training$CompressiveStrength, plot = "pairs")

```

## Question 2, qplot w/o cut


```{r, echo = TRUE}

qplot(row.names(training), CompressiveStrength, data = training)

```


## Question 2, qplot cut2 Cement

```{r, echo = TRUE}

cuts <- cut2(training$Cement, g=5)
plotdata <- cbind(training, cuts)
qplot(row.names(plotdata), CompressiveStrength, data = plotdata, colour = cuts)

```

## Question 2, qplot cut2 BlastFurnaceSlag

```{r, echo = TRUE}

cuts <- cut2(training$BlastFurnaceSlag, g=5)
plotdata <- cbind(training, cuts)
qplot(row.names(plotdata), CompressiveStrength, data = plotdata, colour = cuts)

```

## Question 2, qplot cut2 FlyAsh

```{r, echo = TRUE}

cuts <- cut2(training$FlyAsh, g=5)
plotdata <- cbind(training, cuts)
qplot(row.names(plotdata), CompressiveStrength, data = plotdata, colour = cuts)

```

## Question 2, qplot cut2 Water

```{r, echo = TRUE}

cuts <- cut2(training$Water, g=5)
plotdata <- cbind(training, cuts)
qplot(row.names(plotdata), CompressiveStrength, data = plotdata, colour = cuts)

```

## Question 2, qplot cut2 Superplasticizer

```{r, echo = TRUE}

cuts <- cut2(training$Superplasticizer, g=5)
plotdata <- cbind(training, cuts)
qplot(row.names(plotdata), CompressiveStrength, data = plotdata, colour = cuts)

```

## Question 2, qplot cut2 CoarseAggregate

```{r, echo = TRUE}

cuts <- cut2(training$CoarseAggregate, g=5)
plotdata <- cbind(training, cuts)
qplot(row.names(plotdata), CompressiveStrength, data = plotdata, colour = cuts)

```

## Question 2, qplot cut2 FineAggregate

```{r, echo = TRUE}

cuts <- cut2(training$FineAggregate, g=5)
plotdata <- cbind(training, cuts)
qplot(row.names(plotdata), CompressiveStrength, data = plotdata, colour = cuts)

```

## Question 2, qplot cut2 Age

```{r, echo = TRUE}

cuts <- cut2(training$Age, g=5)
plotdata <- cbind(training, cuts)
qplot(row.names(plotdata), CompressiveStrength, data = plotdata, colour = cuts)

```

Answer: There is a non-random pattern in the plot of the outcome versus index that does not appear to be perfectly explained by any predictor suggesting a variable may be missing.



## Question 3

```{r, echo = TRUE}

library(AppliedPredictiveModeling)
data(concrete)
library(caret)
set.seed(1000)
inTrain = createDataPartition(mixtures$CompressiveStrength, p = 3/4)[[1]]
training = mixtures[ inTrain,]
testing = mixtures[-inTrain,]

```


## Histogram

```{r, echo = TRUE}

ggplot(data=training, aes(training$Superplasticizer)) +
  geom_histogram()

```

Answer: There are values of zero so when you take the log() transform those values will be -Inf.


Question 4

```{r, echo = TRUE}

library(caret)
library(AppliedPredictiveModeling)
set.seed(3433)
data(AlzheimerDisease)
adData = data.frame(diagnosis,predictors)
inTrain = createDataPartition(adData$diagnosis, p = 3/4)[[1]]
training = adData[ inTrain,]
testing = adData[-inTrain,]

```



